TacticToe: Learning the Game of Tactics

### 2018-01-29 tactictoe_jv.tex
minor

\todo{Comparison with hammers in usability}
Many of the state-of-the-art interactive theorem provers (ITPs) such as
  \holfour~\cite{hol4}, \hollight~\cite{Harrison09hollight}, 
  \isabelle~\cite{isabelle} 
  and \coq~\cite{coq-book}
  provide high-level parameterizable tactics for constructing the
  proofs.  Such tactics typically analyze the current goal state and
  assumptions, apply nontrivial proof transformations, which get
  expanded into possibly many basic kernel-level inferences or significant 
  parts of the proof term.
%\marginpar{not Coq}
  In this
  work we develop a tactic-level automation procedure for the \holfour ITP
  which guides selection of the tactics by learning from previous
  proofs.  Instead of relying on translation to first-order automated
  theorem provers (ATPs) as done by the hammer 
  systems~\cite{hammers4qed,tgck-cpp15}, the technique
%\marginpar{too many cites?}
%BlanchetteGKKU16,holyhammer,mizAR40
  directly searches for sequences of tactic applications that lead to
  the ITP proof, thus avoiding the translation and
  proof-reconstruction phases needed by the hammers.

\todo{Rewrite the plan}
  To do this, we \emph{extract and record} tactic invocations from the ITP 
  proofs (Section~\ref{sec:recording}) and
  \emph{build efficient machine learning classifiers} based on such training
  examples (Section~\ref{s:prediction}).  The learned data serves as a 
  guidance for a \emph{Monte Carlo tree search} that explores the different 
  proof paths (Section~\ref{sec:proofsearch}). The result, if
  successful, is a certified human-level proof composed of \holfour
  tactics.  The system is evaluated on a large set of theorems originating
  from \holfour (Section~\ref{s:experiments}), and we show that the performance 
  of the single best \tactictoe strategy exceeds the performance of a hammer 
  system used with a single strategy and a single efficient external prover.

@@

\todoi{remove feature vectors from the paper}
\section{Proof Recording}\label{sec:recording}

@@

Rather than relying on the underlying programming language, we parse the
original theory file containing proof scripts into an
abstract syntax tree (AST).\todo{reason: reading the proof as if you were a 
human ==> better for the machine learning algorithm (learn the right thing)}

 The AST is then transformed into a globalized AST
where every top-level value is made independent from
the context. \todo{some reason why this is needed? easy transfer between 
theories} We 
split the proofs into a series of tactic units and wrap them 
with a recording function.
By replaying the proofs produced by the globalized AST, the wrapper can
record the globalized tactic string with its input goal as training examples in
a database.
The combined effect of these steps on a \holfour proof will be illustrated on 
Example~\ref{ex:running}.

@@

The previously decribed transformation affects each identifier in a theory 
file. However, when we 
create the recording theory file, only proof scripts are replaced by their 
globalized version and the rest of the code is left unchanged. The effect on 
the proof script contained in the running example is shown in 
Example~\ref{ex:global}.

\begin{example} (Globalization)
\begin{lstlisting}[language=SMLSmall,frame=tb]
Tactic.STRIP_TAC 
  infixl0_open boolLib.THEN infixl0_close
Prim_rec.INDUCT_THEN (DB.fetch "list" "list_INDUCT") Tactic.ASSUME_TAC
  infixl0_open boolLib.THEN infixl0_close
Rewrite.ASM_REWRITE_TAC [DB.fetch "list" "MAP", DB.fetch "list" "APPEND"]
\end{lstlisting}
\end{example}

@@

Motivation: 
+ minimal effort to integrate the tactictoe results in proof scripts.

Side effect:
+ could be also used to simplify existing proof scripts.

@@

>>>>

\todoi{Comparison with hammers in usability}
Many of the state-of-the-art interactive theorem provers (ITPs) such as
  \holfour~\cite{hol4}, \hollight~\cite{Harrison09hollight}, 
  \isabelle~\cite{isabelle} 
  and \coq~\cite{coq-book}
  provide high-level parameterizable tactics for constructing the
  proofs.  Such tactics typically analyze the current goal state and
  assumptions, apply nontrivial proof transformations, which get
  expanded into possibly many basic kernel-level inferences or significant 
  parts of the proof term.
%\marginpar{not Coq}
  In this
  work we develop a tactic-level automation procedure for the \holfour ITP
  which guides selection of the tactics by learning from previous
  proofs.  Instead of relying on translation to first-order automated
  theorem provers (ATPs) as done by the hammer 
  systems~\cite{hammers4qed,tgck-cpp15}, the technique
%\marginpar{too many cites?}
%BlanchetteGKKU16,holyhammer,mizAR40
  directly searches for sequences of tactic applications that lead to
  the ITP proof, thus avoiding the translation and
  proof-reconstruction phases needed by the hammers.

\todoi{Rewrite the plan}
  To do this, we \emph{extract and record} tactic invocations from the ITP 
  proofs (Section~\ref{sec:recording}) and
  \emph{build efficient machine learning classifiers} based on such training
  examples (Section~\ref{s:prediction}).  The learned data serves as a 
  guidance for a \emph{Monte Carlo tree search} that explores the different 
  proof paths (Section~\ref{sec:proofsearch}). The result, if
  successful, is a certified human-level proof composed of \holfour
  tactics.  The system is evaluated on a large set of theorems originating
  from \holfour (Section~\ref{s:experiments}), and we show that the performance 
  of the single best \tactictoe strategy exceeds the performance of a hammer 
  system used with a single strategy and a single efficient external prover.

@@

\todoi{Remove feature vectors from the paper}
\section{Proof Recording}\label{sec:recording}

@@

Rather than relying on the underlying programming language, we parse the
original theory file containing proof scripts into an
abstract syntax tree (AST).\todoi{reason: reading the proof as if you were a 
human ==> better for the machine learning algorithm (learn the right thing)}

 The AST is then transformed into a globalized AST
where every top-level value is made independent from
the context. \todoi{some reason why this is needed? easy transfer between 
theories} We 
split the proofs into a series of tactic units and wrap them 
with a recording function.
By replaying the proofs produced by the globalized AST, the wrapper can
record the globalized tactic string with its input goal as training examples in
a database.
The combined effect of these steps on a \holfour proof will be illustrated on 
Example~\ref{ex:running}.

@@

The previously decribed transformation affects each identifier in a theory 
file. However, when creating the recording theory file, only proof scripts are 
replaced by their 
globalized version and the rest of the code is left unchanged. The effect on 
the proof script contained in the running example is shown in 
Example~\ref{ex:global}.

\begin{example}\label{ex:global} (Globalization)
\begin{lstlisting}[language=SMLSmall,frame=tb]
Tactic.STRIP_TAC 
  infixl0_open boolLib.THEN infixl0_close
Prim_rec.INDUCT_THEN (DB.fetch "list" "list_INDUCT") Tactic.ASSUME_TAC
  infixl0_open boolLib.THEN infixl0_close
Rewrite.ASM_REWRITE_TAC [DB.fetch "list" "MAP", DB.fetch "list" "APPEND"]
\end{lstlisting}
\end{example}

@@

\todoi{Motivation: + minimal effort to integrate the tactictoe results in proof 
scripts.}

@@

\todoi{Side effect:
+ could be also used to simplify existing proof scripts. }
